Dark Control:\\ Towards a Unified Account of Default Mode Function\\ by Markov Decision Processes

### 2017-12-08 paper/paper.tex
reread part 5

% , defining $s(t)$, the state at time $t$, as the position in state space.
The process has $memory$ if the next state depends not only on the current state but also on a finite number of past states. Rational probabilistic planning can thus be reformulated as a standard memoryless Markov process by simply expanding the definition of the state $s$ to include experience episodes of the past. This adds the capacity for memory, since then, the next state depends not only on the current state but also on a finite number of past states. This is precisely what Partially Observable MDPs (POMDPs) aim to realize ~\citep{starkweather2017dopamine,o2006making}.
Nevertheless, this mathematical property in POMDPs mostly accounts for implicit $memory$. Since the current paper is focused on both plausibility at the behavioral and neurobiological level, we will address below how our model captures the neurophysiological constraints of the DMN and can also account for the explicit part of $memory$ characteristic for human agents.

@@

\paragraph*{On implicit and explicit memory in the current model.}
Markov Processes are memoryless but this it is mathematically feasible to insert the previous states of such model into the current state. While this may partially account for implicit memory at the behavioral level, it may not explain the underlying implementation at the neurophysiological level nor accounts for the explicit part of memory. Implicit memory-based processing arises in our MDP account of DMN function in three different forms: successive updates of a) the action policy and b) the value function, both being products of the past and c) the deep non-linear links in the hierarchical connections realized by biological neural networks. Indeed, we reason about the brain has deep neural networks that present an implicit form of information compression obtained from life experience. Such memory traces are implicitly stored in the neural machinery and can be retrieve as a form of knowledge during simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}. The neural processes in the hippocampus can be seen as MCMC sampling from a formal statistical level but those memory recall are also the basis for scene construction processes across time-scales such as in the semantic hypothesis \citep{schacter2007remembering,Axelrod2017}.

>>>>

% , defining $s(t)$, the state at time $t$, as the position in state space.
The process has $memory$ if the next state depends not only on the current state but also on a number of past states. Rational probabilistic planning can thus be reformulated as a standard memoryless Markov process by simply expanding the definition of the state $s$ to include experience episodes of the past. This adds the capacity for memory to the model because the next state then depends not only on the current situation but also on previously experienced events, which is the motivation behind Partially Observable MDPs (POMDPs)
\citep{starkweather2017dopamine,o2006making}.
Nevertheless, this mathematical property of POMDPs mostly accounts for implicit $memory$. Since the current paper is concerned with plausibility at the behavioral and neurobiological level, we will address below how our account can accommodate the neurophysiological constraints of the DMN and the explicit $memory$ characteristics of human agents.

@@

\paragraph*{A note on implicit and explicit memory.}
While Markov Processes are usually memoryless, it is mathematically feasible to incorporate the previous states of such model into the current state. This extension may partially account for implicit memory at the behavioral level, but may not explain the underlying neurobiological implementation or accommodate explicit memory. Implicit memory-based processing arises in our MDP account of DMN function in several different forms: successive updates of a) the action policy and the value function, both being products of the past, as well as b) the deep non-linear relationships within the hierarchical connections of biological neural networks. The brain's adaptive synaptic connections can be thought as deep neural networks affording an implicit form of information compression of life experience. Such memory traces are implicitly stored in the neural machinery and can be retrieve as a form of knowledge during simulation of action rather than accessed as a stored explicit representation \citep{pezzulo2011grounding}. c) The neural processes in the hippocampus can be seen as some type of MCMC sampling for memory recall, which can also be a basis for scene construction processes across time-scales \citep{schacter2007remembering,Axelrod2017}.
